Informative Autonomous Sampling of Oceanographic Variables Using Joint Excursion Sets

### 2019-10-15 main_new.tex
Updated figure 10
Updates Jo

\begin{abstract}
 \begin{abstract}
  Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime management. Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provide new opportunities for experimental designs. In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena. Specifically, we define a design criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework. We demonstrate how these provide a description of the boundary uncertainty in temperature and salinity variables, which is used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective by building on myopic and non-myopic strategies. We use simulations to study the properties and differences between state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume. The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.

@@

\item Full numerical ocean models cannot provide accurate results
  online if run on robotic sensing platforms, as onboard computers
  cannot deliver the computational power required, hence statistical
  proxy models of the environment must be used for doing inference
  about where to sample.

\item As a priori information about the state of the ocean is
  uncertain, there is a substantial efficiency gain in reacting to
  information obtained from measurements taken in-situ. Recently
  acquired information can be used to resolve this by assimilating up-to-date data into statistical models that
  can be used to inform decisions on where to sample
  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.

\item For sampling problems related to environmental sensing, the
  number of choices (i.e. locations, trajectories, and candidate
  designs) makes the problem combinatorially large, creating a
  trade-off between optimization (finding the most resource-efficient
  design to collect necessary data), and computability (arriving at a
  solution in reasonable time). To successfully resolve features, this
  trade-off has to be considered in development and practice.

@@

\begin{figure}[!b] 
  \centering 
  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Traditional ocean observation based on
    ship-based sampling has recently been augmented by autonomous and
    collaborative networked operations of robotic vehicles to resolve
    numerous bio-geochemical processes and their interactions. The
    sense-plan-act control approach to autonomous control can reason
    and make decisions for efficient onboard adaptive sampling.
    \ref{fig:nidelven} The interaction of river and ocean creates
    processes that are challenging to map, where the combination of
    statistics and robotics can play a vital role in enabling more
    effective observation practices.}
\label{fig:envir}
\end{figure}

Our focus in this work is towards spatial characterization of a
frontal system specifically, fronts generated by river
plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
the Nidelva river, Trondheim, Norway) where cold freshwater enters
from the river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coreolis force
\citep{coriolis1835memoire} the cold fresh water tends to flow near
land to the east. Depending on the river discharge, tidal effects,
wind, and temperature differences, this boundary often gets
distorted. Prior knowledge about the location and evolution of these
features is therefore highly uncertain, making deterministic planning
challenging.

@@

The Gaussian model is used to represent the bivariate random fields of
temperature and salinity. At location $\bx \in \mathcal{M}$, the prior
bivariate distribution is:
\begin{equation}\label{gp_bar}
  \begin{bmatrix}\xi_T(\bx) \\
    \xi_S(\bx) \end{bmatrix}
 \sim N \left( 
\begin{bmatrix} \mu_{T}(\bx)\\
\mu_{S}(\bx)
\end{bmatrix},\begin{bmatrix}
\sigma_{T}^2(\bx) & \sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  \\
\sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx) 
\end{bmatrix}
\right),
\end{equation}
where the notation refers to Gaussian (i.e. normal) distributed
variables specified by the mean vector and covariance matrix. For the
application of a river plume, the model parameters are specified from
a short preliminary survey. The means vary with location, capturing
the variability near the river mouth, while the variance parameters
and correlation between salinity and temperature are assumed to be
constant for all locations. %and within each water mass.

For the spatial dependence among variables, we assume a separable
correlation function with the same decay for salinity and temperature,
which is consistent with expectations for a river plume where the
decay is related to common water masses (Section
\ref{sec:case_study}). In the resulting function, we have:
\begin{equation}\label{gp_corr}
\mbox{Cov}(\xi_i(\bx),\xi_j(\bx')) = \rho(\bx,\bx') \sigma_i \sigma_j (\gamma +(1-\gamma )\delta_{ij}), \\
\end{equation}
where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and
$\delta_{ij}=0$ otherwise. Moreover, we assume stationary isotropic
random fields a priori where the correlation function $\rho(\bx,\bx')$
solely depends on $\bx$ and $\bx'$ via Euclidean distance
$h=\sqrt{||\bx-\bx'||^2}$. With data and prior knowledge, one could
possibly fit and estimate parameters of non-stationary or
non-isotropic covariance functions and more complex multivariate
spatial covariance functions
\citep{gneiting2010matern,genton2015cross}; however, that is left for
future work.

@@

Table \ref{tab:sim_rhoab} shows the EPs and the associated prior Bernoulli
variance (second row) for the examples indicated in
Fig. \ref{illus_bivarDens}. The EPs increase with the correlation as
there is a strong tendency to have concurrently low temperature and
salinity. The Bernoulli variance is similarly large for high
correlations. EPs and Bernoulli variances are the same for standard
deviation $1$ or $2$, which implies that high variability in
temperature and salinity is not captured in the $p(1-p)$ expression.

@@

We use a bivariate GP model for temperature and salinity, where we
specify the prior mean:
\begin{equation}\label{m}
    \bmu(\bx)=E 
    \begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix}=\begin{bmatrix} \mu_{T}(\bx)\\
\mu_{S}(\bx)
\end{bmatrix} 
= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{West}} \\
\beta_{S,0} + \beta_{S,1} x_{\mbox{West}}
\end{bmatrix}.
\end{equation}
This situation
mimics that of a river mouth opening from the south, with the water
masses pulled to the east. The regression parameters are  $\beta_{T,1}=0.065$ and
$\beta_{S,1}=0.1$ for the slopes with west coordinate, and $\beta_{T,0}=5.8$
and $\beta_{S,0}=29.0$ for the intercept at a chosen reference map coordinate.

@@

The covariance is assumed to be stationary and separable for the two variables. 
The $2 \times 2$ pointwise covariance matrix is set to:
\begin{equation}\label{v0}
\bSigma(\bx)=\mbox{Var} 
\begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix}=
\begin{bmatrix}
0.25^2 & 0.6 \cdot 0.25^2 \\
0.6 \cdot 0.25^2 & 0.25^2
\end{bmatrix}.
\end{equation}
The spatial correlation is of a Matern type \kc{cite or define}:
\begin{equation}\label{v}
\small
\mbox{Corr} 
\left(
\begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix},
    \begin{bmatrix}
    \xi_T(\bx') \\
    \xi_S(\bx') 
    \end{bmatrix}
    \right)
    = \begin{bmatrix}
1 & 0.6  \\
0.6  & 1
\end{bmatrix}(1+\phi h)\exp (-\phi h),
\end{equation}
where $h=\sqrt{||\bx-\bx'||^2}$ and $\phi=0.3$ indicates an effective correlation range of about $1200$ m. 
%Figure \ref{fig:stat_design} shows the contour lines for the EP for the reference. We notice the trend of increasing salinity and temperature to the west.  
One realization of the random fields representing salinity and
temperature is shown in Fig. \ref{fig:true_temp} and
\ref{fig:true_sal}. The true ES from this prior realization is shown
in Fig. \ref{fig:ESet}.

@@

For a fixed survey length, a closed form expression for the expected IBV is available as in Eq. \eqref{sur}. However, for the sequential approaches this is not the
case. For comparison the properties are therefore evaluated using
Monte Carlo sampling over several replicates of realizations from the
prior while conducting simulated sequential surveys for each
one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
gathered along the north-south survey lines in these figures. In the
Monte Carlo replicates, such results are repeatedly computed to
approximate the expected IBV. We also compare predictive performance
measured by root mean square error (RMSE) for temperature and salinity
estimates as also the variance reduction in these two variables. It is
important to note that the objective function used by the agent is
focused on reducing the expected IBV, but we nevertheless expect that
we will achieve good predictive performance for criteria such as RMSE
as well. Another non-statistical criteria that is important for
practical purposes is the computational time needed for the strategy,
as this will impact the performance for an embedded system.

@@

In Fig. \ref{fig:route_choices}, the realized sampling paths for each
of the sequential schemes and static designs are shown. The
\textit{naive} strategy often gets stuck in the southern part of the
domain because it is too focused on the probabilities near $0.5$. The
\textit{myopic} strategy covers a wider domain than the naive or
look-ahead. There are several reasons for this. First, a greedy
approach will tend to put more emphasis on promising locations close to the
agent, which may lead away from the centre. Second, as the agent
evaluates the impact of locations further away (look-ahead) where
assimilated data has less predictive power, the GP prior model (which
is centered here) will act to restrict paths deviating from the
central zone.

We studied the sensitivity of the results by modifying the input parameters to different prior correlations between temperature and salinity, standard deviations, and spatial correlation range.  In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}. The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small prior standard deviation or very large spatial correlation range. The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range. We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the effects are not much worse in terms of IBV when only temperature data are available. In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data. However, it seems that having temperature data alone does a substantially worse job in terms of explained variance

@@

The tables below explore the sensitivity of the result to different
prior correlation values between temperature measured onboard the AUV,
temperature and salinity, prior standard deviations, and spatial
correlations (see table rows). We only compare results at the end of
each survey. The expected ES variances are shown in Table
\ref{tab:sim_res_ev}, results of RMSE are in Table
\ref{tab:sim_res_rmse}, and explained variance is in Table
\ref{tab:sim_res_r2}. In all runs, both the \textit{myopic} and
\textit{look-ahead} strategies perform the best in terms of expected
variance in the ES. The \textit{myopic} strategy
performs slightly better than the \textit{look-ahead}. Part of the
reason for this discrepancy can be traced to the act of pruning some
of the paths for the \textit{look-ahead} strategy. The \textit{static\_north} design is clearly
better than the other two static designs, the reason being that the
others do not focus on the areas where the ES is really uncertain
(centre locations are the most uncertain). By measuring temperature
alone, the expected variances in the ES are only slightly higher than
the combined temperature-salinity estimate. This occurs because the
correlation of $0.6$ means that temperature observations provide
information about salinity as well. For the other predictive measures,
\textit{myopic} and \textit{look-ahead} strategies do well, clearly
beating the \textit{naive} strategy.

@@

\subsection{Model Specification}
\label{sec:exp_modeling}
The statistical model parameters were specified based on a short
preliminary survey where the AUV made an initial transect to determine
environmental conditions and correlation structure. Based on the
initial data, the trend parameters were estimated by linear
regression, where both temperature and salinity are assumed to
increase linearly, going west from the river mouth. Next, the
residuals from the regression analysis were analyzed to study the fit
of the GP model and to specify the covariance parameters.

@@

Fig. \ref{fig:parest} shows diagnostic plots of this
analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the westerly trends in both salinity and
temperature are subtracted from the data. This scatter-plot of joint
residuals indicates larger variability in salinity than in
temperature, and a positive correlation ($0.5$) between the two
variables. Based on the fitted bivariate Gaussian model (ellipses in
Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
the residuals, and if the model is adequate they should be
approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
the empirical cumulative distribution function (CDF) of these
quadratic forms (solid) together with the theoretical CDF of the
$\chi^2_2$ distribution. The modeled and theoretical curves are very
similar, which indicates that the Gaussian model fits reasonably
well. Even though there appears to be some clustering in both
Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate
diagnostic plots look reasonable and justify a Gaussian
model. Fig. \ref{fig:parest_c} shows the empirical variogram of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m.

@@

The final predictions of the front location given in
Fig. \ref{fig:res1} and \ref{fig:res2} as a dashed line, correspond
with one another, yielding a picture of the front gradually bending
off toward the north east. The amount of exploration done by Survey 1
is greater than Survey 2. In Survey 1, the AUV obtained more detail by
going north, while Survey 2, coming close to the survey area borders
in the south-western corner, obtained a poorer understanding of the
northern parts. A longer planning horizon would likely identify and
discourage such choices.

%\begin{figure}[!h] 
% \centering 
%\includegraphics[width=0.48\textwidth]{Figures/envir_ocean.pdf}
%\caption{Ocean observation is moving away from %single-ship sampling
%towards more collaborative networked operations in %order to resolve
%the numerous processes and their interaction.} %\label{fig:envir}
%\end{figure}

\section{Closing remarks}\label{sec:concl_disc}

>>>>

\begin{abstract}

  Improving and optimizing oceanographic sampling is a crucial task for marine science and maritime management. Faced with limited resources to understand processes in the water-column, the combination of statistics and autonomous robotics provides new opportunities for experimental designs. In this work we develop methods for efficient spatial sampling applied to the mapping of coastal processes by providing informative descriptions of spatial characteristics of ocean phenomena. Specifically, we define a design criterion based on reducing uncertainty in the joint excursions of vector-valued Gaussian random fields, and derive tractable expressions for the expected Bernoulli variance reduction in such a framework. We demonstrate how this criterion gives a description of the boundary uncertainty in temperature and salinity variables, which is used to prioritize sampling efforts at locations that are ambiguous, making exploration more effective by building on myopic and non-myopic strategies. We use simulations to study the properties of methods and to compare them with state-of-the-art approaches, followed by results from field deployments with an autonomous underwater vehicle as part of a case study mapping the boundary of a river plume. The results demonstrate the potential of combining statistical methods and robotic platforms to effectively inform and execute data-driven environmental sampling.

@@

\item Full numerical ocean models cannot provide accurate results
  online if run on robotic sensing platforms, as onboard computers
  cannot deliver the computational power required. Hence statistical
  proxy models of the environment must be used for learning where to sample.

\item With limited available information about the state of the ocean, there is substantial value in reacting to
  information obtained from measurements taken in-situ. This
  acquired information must be assimilated into statistical models that
  can be used to inform decisions on where to sample
  sequentially.%\emph{in-situ}; this is usually referred to as the \emph{adaptivity gap} \citep{Krause2008phd}.

\item For sampling problems related to environmental sensing, the
  number of choices (i.e. locations, trajectories, and candidate
  designs) is enormous, creating a
  trade-off between optimization (finding the most resource-efficient
  design to collect necessary data) and computability (arriving at a
  solution in reasonable time). To successfully resolve features, this
  trade-off has to be considered in development and practice.

@@

\begin{figure}[!h] 
  \centering 
  \subfigure[Illustration of an ocean sensing network and the sense-plan-act control methodology.]{\includegraphics[width =
    0.49\textwidth]{Figures/envir2.pdf}\label{fig:envir1}}
  \hfill
  \subfigure[Frontal patterns off of the Nidelva river, Trondheim, Norway.]{\includegraphics[width =
    0.49\textwidth]{Figures/river_proccess.pdf}\label{fig:nidelven}}
  \caption{\ref{fig:envir1} Traditional ocean observation based on
    ship-based sampling has recently been augmented by autonomous and
    collaborative networked operations of robotic vehicles to resolve
    numerous bio-geochemical processes and their interactions. The
    sense-plan-act control approach to autonomous control can reason
    and make decisions for efficient onboard adaptive sampling.
    \ref{fig:nidelven} The interaction of river and ocean creates
    processes that are challenging to map, where the combination of
    statistics and robotics can play a vital role in enabling more
    effective observation practices.}
\label{fig:envir}
\end{figure}

Our focus in this work is towards spatial characterization of a
frontal system specifically, fronts generated by river
plumes. Fig. \ref{fig:nidelven} shows the survey area (the outlet of
the Nidelva river, Trondheim, Norway) where cold freshwater enters
from the river, creating a strong gradient in both temperature and
salinity. Because of the local topography and the Coreolis force
\citep{coriolis1835memoire} the cold fresh water tends to flow near
land to the east. Depending on the river discharge, tidal effects,
wind, and temperature differences, this boundary often gets
distorted. Initial knowledge about the location and evolution of these
features is therefore highly uncertain, making deterministic planning
challenging.

@@

The Gaussian model is used to represent the bivariate random fields of
temperature and salinity. At location $\bx \in \mathcal{M}$, the
bivariate distribution is:
\begin{equation}\label{gp_bar}
  \begin{bmatrix}\xi_T(\bx) \\
    \xi_S(\bx) \end{bmatrix}
 \sim N \left( 
\begin{bmatrix} \mu_{T}(\bx)\\
\mu_{S}(\bx)
\end{bmatrix},\begin{bmatrix}
\sigma_{T}^2(\bx) & \sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  \\
\sigma_{T}(\bx) \sigma_S(\bx) \gamma(\bx)  & \sigma_{S}^2(\bx) 
\end{bmatrix}
\right),
\end{equation}
where the notation refers to Gaussian (i.e. normal) distributed
variables specified by the mean vector and covariance matrix. For the
application of a river plume, the model parameters are specified from
a short preliminary survey. The means vary with location, capturing
the variability near the river mouth, while the variance parameters
and correlation between salinity and temperature are assumed to be
constant for all locations. %and within each water mass.

For the spatial dependence among variables, we assume a separable
correlation function with the same decay for salinity and temperature,
which is consistent with expectations for a river plume where the
decay is related to common water masses (Section
\ref{sec:case_study}). In the resulting function, we have:
\begin{equation}\label{gp_corr}
\mbox{Cov}(\xi_i(\bx),\xi_j(\bx')) = \rho(\bx,\bx') \sigma_i \sigma_j (\gamma +(1-\gamma )\delta_{ij}), \\
\end{equation}
where for $i,j \in {T,S}$ and $\delta_{ij}=1$ if $i=j$, and
$\delta_{ij}=0$ otherwise. Moreover, we assume stationary isotropic
random fields where the correlation function $\rho(\bx,\bx')$
solely depends on $\bx$ and $\bx'$ via Euclidean distance
$h=\sqrt{||\bx-\bx'||^2}$. With data and prior knowledge, one could
possibly fit and estimate parameters of non-stationary or
non-isotropic covariance functions and more complex multivariate
spatial covariance functions
\citep{gneiting2010matern,genton2015cross}; however, that is left for
future work.

@@

Table \ref{tab:sim_rhoab} shows the initial EPs and the associated Bernoulli
variance (second row) for the examples indicated in
Fig. \ref{illus_bivarDens}. The EPs increase with the correlation as
there is a strong tendency to have concurrently low temperature and
salinity. The Bernoulli variance is similarly large for high
correlations. EPs and Bernoulli variances are the same for standard
deviation $1$ or $2$, which implies that high variability in
temperature and salinity is not captured in the $p(1-p)$ expression.

@@

We use a bivariate GP model for temperature and salinity, where we
specify the mean:
\begin{equation}\label{m}
    \bmu(\bx)=E 
    \begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix}=\begin{bmatrix} \mu_{T}(\bx)\\
\mu_{S}(\bx)
\end{bmatrix} 
= \begin{bmatrix} \beta_{T,0} + \beta_{T,1} x_{\mbox{West}} \\
\beta_{S,0} + \beta_{S,1} x_{\mbox{West}}
\end{bmatrix}.
\end{equation}
This situation
mimics that of a river mouth opening from the south, with the water
masses pulled to the east. The regression parameters are  $\beta_{T,1}=0.065$ and
$\beta_{S,1}=0.1$ for the slopes with west coordinate, and $\beta_{T,0}=5.8$
and $\beta_{S,0}=29.0$ for the intercept at a chosen reference map coordinate.

@@

The covariance is assumed to be stationary and separable for the two variables. 
The $2 \times 2$ pointwise covariance matrix is set to:
\begin{equation}\label{v0}
\bSigma(\bx)=\mbox{Var} 
\begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix}=
\begin{bmatrix}
0.25^2 & 0.6 \cdot 0.25^2 \\
0.6 \cdot 0.25^2 & 0.25^2
\end{bmatrix}.
\end{equation}
The spatial correlation is of a Matern type \kc{cite or define}:
\begin{equation}\label{v}
\small
\mbox{Corr} 
\left(
\begin{bmatrix}
    \xi_T(\bx) \\
    \xi_S(\bx) 
    \end{bmatrix},
    \begin{bmatrix}
    \xi_T(\bx') \\
    \xi_S(\bx') 
    \end{bmatrix}
    \right)
    = \begin{bmatrix}
1 & 0.6  \\
0.6  & 1
\end{bmatrix}(1+\phi h)\exp (-\phi h),
\end{equation}
where $h=\sqrt{||\bx-\bx'||^2}$ and $\phi=0.3$ indicates an effective correlation range of about $1200$ m. 
%Figure \ref{fig:stat_design} shows the contour lines for the EP for the reference. We notice the trend of increasing salinity and temperature to the west.  
One realization of the random fields representing salinity and
temperature is shown in Fig. \ref{fig:true_temp} and
\ref{fig:true_sal}. The true ES is a realization from this model as shown
in Fig. \ref{fig:ESet}.

@@

For a fixed survey length, a closed form expression for the expected IBV is available as in Eq. \eqref{sur}. However, for the sequential approaches this is not the
case. For comparison the properties are therefore evaluated using
Monte Carlo sampling over several replicates of realizations from the model while conducting simulated sequential surveys for each
one. Fig. \ref{fig:Eprob} shows the conditional EP, given data
gathered along the north-south survey lines in these figures. In the
Monte Carlo replicates, such results are repeatedly computed to
approximate the expected IBV. We also compare predictive performance
measured by root mean square error (RMSE) for temperature and salinity
estimates as also the variance reduction in these two variables. It is
important to note that the objective function used by the agent is
focused on reducing the expected IBV, but we nevertheless expect that
we will achieve good predictive performance for criteria such as RMSE
as well. Another non-statistical criteria that is important for
practical purposes is the computational time needed for the strategy,
as this will impact the performance for an embedded system.

@@

In Fig. \ref{fig:route_choices}, the realized sampling paths for each
of the sequential schemes and static designs are shown. The
\textit{naive} strategy often gets stuck in the southern part of the
domain because it is too focused on the probabilities near $0.5$. The
\textit{myopic} strategy covers a wider domain than the naive or
look-ahead. There are several reasons for this. First, a greedy
approach will tend to put more emphasis on promising locations close to the
agent, which may lead away from the centre. Second, as the agent
evaluates the impact of locations further away (look-ahead) where
assimilated data has less predictive power, the GP model (which
is centered here) will act to restrict paths deviating from the
central zone.

We studied the sensitivity of the results by modifying the input parameters to have different correlations between temperature and salinity, standard deviations, and spatial correlation range.  In all runs,  the \textit{myopic} and \textit{look-ahead} strategies perform the best in terms of IBV, and much better than \textit{naive}. The \textit{look-ahead} strategy seems to be substantially better than the \textit{myopic} only for very small initial standard deviation or very large spatial correlation range. The \textit{static\_north} design continues to be the best static design for IBV, while \textit{static\_zigzag} is the best design for the other predictive performance measures, especially so with large spatial correlation range. We also ran simulation studies with only temperature data, and for realistic correlation levels between temperature and salinity the IBV results are not much worse when only temperature data are available. In addition to the comparison made in Table \ref{tab:sim_rhoab}, the current setting includes spatial correlation and this likely reduces the additional influence of having bivariate data. However, it seems that having temperature data alone does a substantially worse job in terms of explained variance

@@

The tables below explore the sensitivity of the result to different input correlation values between temperature and salinity measured onboard the AUV, standard deviations, and spatial
correlations (see table rows). We only compare results at the end of
each survey. The expected ES variances are shown in Table
\ref{tab:sim_res_ev}, results of RMSE are in Table
\ref{tab:sim_res_rmse}, and explained variance is in Table
\ref{tab:sim_res_r2}. In all runs, both the \textit{myopic} and
\textit{look-ahead} strategies perform the best in terms of expected
variance in the ES. The \textit{myopic} strategy
performs slightly better than the \textit{look-ahead}. Part of the
reason for this discrepancy can be traced to the act of pruning some
of the paths for the \textit{look-ahead} strategy. The \textit{static\_north} design is clearly
better than the other two static designs, the reason being that the
others do not focus on the areas where the ES is really uncertain
(centre locations are the most uncertain). By measuring temperature
alone, the expected variances in the ES are only slightly higher than
the combined temperature-salinity estimate. This occurs because the
correlation of $0.6$ means that temperature observations provide
information about salinity as well. For the other predictive measures,
\textit{myopic} and \textit{look-ahead} strategies do well, clearly
beating the \textit{naive} strategy.

@@

\subsection{Model Specification}
\label{sec:exp_modeling}
The statistical model parameters were specified based on a short
preliminary survey where the AUV made an initial transect to determine the trends in environmental conditions and correlation structures. Based on the
initial data, the trend parameters were estimated by linear
regression, where both temperature and salinity are assumed to
increase linearly, going west from the river mouth. Next, the
residuals from the regression analysis were analyzed to study the fit
of the GP model and to specify the covariance parameters.

@@

Fig. \ref{fig:parest} summarizes diagnostic plots of this
analysis. Fig. \ref{fig:parest_a} shows a cross-plot of temperature
and salinity residuals after the westerly trends in both salinity and
temperature are subtracted from the data. This scatter-plot of joint
residuals indicates larger variability in salinity than in
temperature, and a positive correlation ($0.5$) between the two
variables. Based on the fitted bivariate Gaussian model (ellipses in
Fig. \ref{fig:parest_a}), we can compute the modeled quadratic form of
the residuals, and if the model is adequate they should be
approximately $\chi^2_2$ distributed. Fig. \ref{fig:parest_b} shows
the empirical cumulative distribution function (CDF) of these
quadratic forms (solid) together with the theoretical CDF of the
$\chi^2_2$ distribution. The modeled and theoretical curves are very
similar, which indicates that the Gaussian model fits reasonably
well. Even though there appears to be some clustering in both
Fig. \ref{fig:parest_a} and \ref{fig:parest_b}, the bivariate
diagnostic plots look reasonable and justify a Gaussian
model. Fig. \ref{fig:parest_c} shows the empirical variogram of the
scaled residuals for temperature and salinity. The decay is similar
for the two, and seems to be negligible after about $150$ m.

@@

The final predictions of the front location, represented by conditional EPs in
Fig. \ref{fig:res1} and \ref{fig:res2} as dashed lines, correspond
with one another. In both surveys they yield a picture of the front being a bit to the west in the southern parts of the domain and gradually bending off toward the north east. The amount of exploration done by Survey 1
is greater than Survey 2. In Survey 1, the AUV obtained more detail by
going north from waypoint 5, while Survey 2, coming close to the survey area borders
in the south-western corner, obtained a poorer understanding of the
northern parts. A look-ahead strategy might identify and
discourage such choices.

%\begin{figure}[!h] 
% \centering 
%\includegraphics[width=0.48\textwidth]{Figures/envir_ocean.pdf}
%\caption{Ocean observation is moving away from %single-ship sampling
%towards more collaborative networked operations in %order to resolve
%the numerous processes and their interaction.} %\label{fig:envir}
%\end{figure}
\newpage
\section{Closing remarks}\label{sec:concl_disc}
